---
title: "Compare ArcMap and akgfmaps grids in the EBS and NBS"
date: "3/16/2022"
output: word_document
---

## Compare ArcMap and akgfmaps grids in the EBS and NBS

The purpose of this analysis is to compare EBS and NBS grids generated using the fishnet tool and area calculator in ArcMap to grids generated using the akgfmaps::make_2d_grid function. The goal for the akgfmaps::make_2d_grid function is to reproduce results of grid production methods in ArcMap.

Grids are used as interpolation grids in VAST.

```{r load, include=FALSE}
library(akgfmaps) # Version >= 1.8.0
```

```{r setup, message=FALSE, warning=FALSE, echo=TRUE}
# Load Jason's csv output
arcmap_ebs <- read.csv(file = here::here("data", "VAST_raster_EBS_2022.csv"), stringsAsFactors = FALSE)
arcmap_nbs <- read.csv(file = here::here("data", "VAST_raster_NBS_2022.csv"), stringsAsFactors = FALSE)

# Setup connection to spatial products gdb
dsn <- here::here("C:/Users/sean.rohan/Work/afsc/WIP", "Bering_Sea_Spatial_Products_2022.gdb")

# Setup grid using dimensions from Jason's template
vast_bbox = c(xmin = -2846314.206900,
         ymin = 251495.775400,
         xmax = 2128157.793100,
         ymax = 2859111.775400)

res <- c(3704, 3704)
```


## Make grids using akgfmaps::make_2d_grid

```{r make_2d_grid, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
ebs_grid <- akgfmaps:::make_2d_grid(obj = sf::st_read(dsn = dsn,
                                                      layer = "EBS_strata_Conner2022"),
                                    resolution = res,
                                    bbox = vast_bbox,
                                    output_type = "point")

nbs_grid <- akgfmaps:::make_2d_grid(obj = sf::st_read(dsn = dsn,
                                                      layer = "NBS_strata_Conner2022"),
                                    resolution = res,
                                    bbox = vast_bbox,
                                    output_type = "point")
```


## Total stratum areas

Calculate total stratum areas from ArcMap and akgfmaps for the EBS and NBS.

```{r calculate_areas, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}
stratum_total_arcmap_ebs <- arcmap_ebs %>%
  dplyr::group_by(Stratum) %>%
  dplyr::summarise(Stratum_Area = sum(Area_KM2))

stratum_total_akgfmaps_ebs <- ebs_grid %>%
  as.data.frame() %>%
  dplyr::group_by(Stratum) %>%
  dplyr::summarise(Stratum_Area = sum(as.numeric(AREA)/1e6))

stratum_total_arcmap_nbs <- arcmap_nbs %>%
  as.data.frame() %>%
  dplyr::group_by(Stratum) %>%
  dplyr::summarise(Stratum_Area = sum(as.numeric(Area_KM2)))

stratum_total_akgfmaps_nbs <- nbs_grid %>%
  as.data.frame() %>%
  dplyr::group_by(Stratum) %>%
  dplyr::summarise(Stratum_Area = sum(as.numeric(AREA)/1e6))
```


## Check for differences in number of cells


```{r nrow_ebs}
nrow(ebs_grid)
nrow(arcmap_ebs)

table(ebs_grid$Stratum)
table(arcmap_ebs$Stratum)
```

```{r nrow_nbs}
nrow(nbs_grid)
nrow(arcmap_nbs)

table(nbs_grid$Stratum)
table(arcmap_nbs$Stratum)
```

EBS and NBS grids from akgfmaps and ArcMap have the same number of cells.


## Check for differences in total stratum area

```{r stratum area}
(stratum_total_arcmap_ebs$Stratum_Area - stratum_total_akgfmaps_ebs$Stratum_Area)
stratum_total_arcmap_nbs$Stratum_Area - stratum_total_akgfmaps_nbs$Stratum_Area
```

EBS differences in total stratum area on the order of 1.2-48.8 m^2. NBS has no differences in total stratum area.


## Example: Transform to WGS84

Transforming the grid points to WGS84 (EPSG:4326)

```{r ex_transform}
sf::st_transform(ebs_grid, crs = "EPSG:4326")
```
